\section{Future Work}

Despite the promising results obtained in this study, the scope of the evaluation
has been limited by time and hardware availability. Consequently, a number of
important directions remain open for future work. Addressing these aspects could
provide a more comprehensive understanding of the applicability and robustness
of power and energy monitoring approaches on AMD processors.

First, a broader evaluation across a larger set of AMD \gls{CPU} families should
be carried out. While this work focused on desktop-class Zen~2 processors (e.g.,
Matisse), server-grade AMD EPYC \gls{CPU}s are increasingly deployed in modern
clusters and data centers. Investigating the behavior of energy and power
counters, scaling properties, and reliability of readings across different EPYC
generations (e.g., Rome, Milan, Genoa) would be of particular relevance for
high-performance computing environments and cloud infrastructures
\parencite{hackenberg2019epyc}.

Second, a systematic comparison between the available AMD software stacks could
enrich the findings. For instance, AMD’s official ecosystem tools such as the
\textit{e-SMI} library provide access to low-level telemetry and power
management functionality, but their coverage and stability differ from both
\gls{RAPL} and the EMA extensions considered in this work. Comparative studies
between \textit{e-SMI}, \gls{RAPL}, and EMA-based plugins could clarify
trade-offs in accuracy, portability, and performance overhead
\parencite{amd2022esmi}.

In addition to vendor-provided libraries, the community maintains several Linux
kernel drivers and user-space utilities, including the well-known
\textit{Zenpower} driver. Evaluating these open-source solutions side by side
with vendor tools would not only provide insights into their accuracy and
usability but also highlight the potential for cross-validation of results in
research and operational settings.

Finally, although this study focused exclusively on Linux, extending the
methodology to other operating systems could be beneficial. In particular,
investigating the availability of similar low-level counters and power
management interfaces under Windows could reveal whether analogous monitoring
approaches can be generalized beyond Unix-like environments. While the
practicality of this direction depends on driver and \gls{API} support, it may
open opportunities for cross-platform benchmarking and validation.

In summary, future work should prioritize a more diverse hardware base,
systematic cross-tool comparisons, and potentially cross-platform studies. Taken
together, these directions would contribute to a more complete and reliable
assessment of energy monitoring capabilities on modern AMD processors.
