\section{Literature Review}

The question of measuring and analyzing \gls{CPU} energy consumption has
attracted increasing attention in recent years. Several studies have
focused on the use of the \gls{RAPL} interface,
which remains the dominant software-accessible mechanism for this
purpose, while comparatively little research has investigated
AMD-specific interfaces such as \texttt{ryzen\_smu}.

The following subsection summarizes the most relevant works and
highlights the persisting gaps that motivate the present research.

\subsection{Related Work}

\textcite{EnergyEffZen2} provide an in-depth analysis of energy efficiency
characteristics of the AMD Zen 2 architecture. Their evaluation employs the
\gls{RAPL} interface in combination with performance counters and low-level
profiling tools. The focus is primarily on architectural behavior,
instruction-level performance, and the efficiency of specific
micro-architectural features rather than on the comparative accuracy of
energy measurement techniques themselves.

Although the work demonstrates how \gls{RAPL} can be integrated into
architectural studies, it does not investigate the validity of
alternative AMD-specific \gls{SMU} interfaces.

A complementary line of research is represented by \textcite{RAPL_Strategies},
who systematically examine strategies for employing \gls{RAPL} to measure
energy consumption in high-performance computing workflows. Their study
emphasizes practical aspects of where and how measurements can be integrated
into execution pipelines (e.g., at the application code level, through external
wrappers, or via plugin mechanisms). The authors further explore the relationship
between input data size and observed energy consumption, offering insights
that are valuable for workload-level optimization. However, the study is limited
to methodological strategies around \gls{RAPL} and does not assess the
comparative strengths or weaknesses of alternative measurement interfaces.

Finally, \textcite{SoftMeasComp} present a broad comparative analysis of
software-based energy measurement techniques across Intel and AMD
processors. Their work offers a particularly comprehensive overview of
the reliability and limitations of \gls{RAPL}, with detailed discussion of
operating system integration and implementation differences. Importantly,
while AMD processors are included in the dataset, the analysis remains
confined to \gls{RAPL}-based measurements, without incorporating \gls{SMU}
interfaces such as \texttt{ryzen\_smu}. 

Across all three studies, several recurring issues with \gls{RAPL} are
identified, including undocumented or poorly defined energy domains,
potential double-counting in hierarchical aggregation, register
overflows, and the well-known susceptibility of \gls{RAPL} to side-channel
attacks \parencite[see e.g.][]{ RAPL_SideChannel_Spy}.
These limitations underscore the need for
systematic evaluation of alternative AMD-specific mechanisms, which this
work seeks to address.

Since all the studies mentioned above used \gls{RAPL} extensively, they where also
used to infer useful technical information, insights and references, allowing
the author to extend the contextual understanding of the interface.

\subsection{Supporting Documentation}

In addition to research articles, several forms of supporting
documentation were instrumental in the preparation and execution of
this study. These resources provided both the practical foundation for
system configuration and the technical background required for working
with AMD-specific interfaces, Linux monitoring tools and kernel drivers.

The Red Hat documentation and ArchWiki were frequently consulted during
system setup and software preparation. These resources were essential in
the identification and configuration of key tools, including hardware
monitoring utilities, processor and/or mainboard topology inspection
tools, kernel modules and drivers, process management mechanisms, etc.
Beyond initial configuration, both sources served as references for
troubleshooting and verification of software and hardware functionality for
target system, thereby contributing directly to the reliability of the
experimental environment and setup.
\parencite{RedHat_Docs, ArchWiki}

For AMD-specific mechanisms, official vendor documentation proved less
useful due to limited accessibility and fragmented nature. In practice,
one the most valuable reference was the open-source \texttt{ryzen\_smu} kernel
driver repository \parencite{RyzenSMU_GitHub}. Alongside with the source code
it contains fairly detailed documentation in its accompanying README.
To overcome some challenges during methodology development the
\texttt{ryzen\_monitor} repository \footcite{RyzenMonitor_GitHub} was also used
as a source of some implementation and usage examples of the driver. Direct
inspection of the mentioned material provided critical insights into available
usage variants, functions, register mappings, and constraints, which were essential
for the implementation of the required/missing software parts required for
successful completions of this research.

Intel’s official developer manuals, which are widely regarded
as a reference standard in terms of accessibility and comprehensiveness,
were used primarily as a comparative benchmark \parencite{Intel_SDM}. Their
structured presentation of hardware interfaces and architectural features
highlighted the relative scarcity and fragmentation of AMD-specific documentation.

A complementary source of technical insight is the informal GitHub
repository \\\emph{plops/ryzen\_managment\_linux}, which provides annotated
explanations of the low-level behavior of the \texttt{ryzen\_smu} driver
interface. In particular, the author highlights that the SMU driver is
\emph{on-demand} rather than polling, and implements an internal throttling
mechanism—limiting \gls{SMU} refreshes to at most 1 kHz—even if userspace attempts
to poll faster \parencite{ryzen_managment_linux}. While this project is not a
formal documentation, it helps to understand practical limitations of
the \texttt{pm\_table} interface and guides decisions about sampling
frequency and overhead optimization during plugin implementation.

\subsubsection{Official Telemetry Sources}

This study relies on a combination of vendor documentation, official
tools, and in-band libraries to establish a reproducible and verifiable
telemetry baseline for AMD hardware. These sources provide the
architectural context, recommended system tuning, and \gls{API}/driver-level
details required to implement and validate measurement adapters.

\begin{itemize}
  \item \textbf{AMD architecture and power whitepapers.} Official AMD
  whitepapers describe the System Management Unit (SMU), per-die agents,
  and power domains on EPYC/Ryzen platforms. These documents are used to
  define measurement scope (which hardware domains are exposed), to
  interpret telemetry semantics, and to justify the comparability
  between SMU-derived and RAPL-derived metrics
  \parencite{AMD_EPYC_9004_8004_PowerMgmt_2024,
  AMD_EPYC_5th_Arch_2024}.

  \item \textbf{In-band libraries and E-SMI.} Libraries such as E-SMI
  (esmi\_ib\_library) expose programmatic access to SMU/EPYC telemetry.
  These \gls{API}s were consulted for implementation patterns and (where
  applicable) used to validate readings obtained through the
  \texttt{ryzen\_smu} driver \parencite{AMD_esmi_ib_library}.

  \item \textbf{Tuning and platform configuration guides.} Official
    platform tuning guides and vendor recommendations (OS tuning, \gls{CPU}
  governors, BIOS settings) were followed to ensure reproducible test
  conditions and to avoid systemic distortions of power measurements
  \parencite{AMD_RHEL_Tuning_9004_2024}.

  \item \textbf{Security and limitations.} Research on power-reporting
  side channels motivates conservative interpretation of high-resolution
  telemetry and informs safety/privacy mitigations when exposing raw
  power or energy counters \parencite{PwrLeak_2023}.
\end{itemize}

The combined use of these official documents and tools supports: (i)
correct identification of measurement domains; (ii) reproducible
platform configuration; (iii) selection of implementation approaches fo
the \gls{SMU}-to-energy adapter; and (iv) principled discussion of the limitations
and security considerations of telemetry sources.
