\section*{Abstract}

With the increasing demand for computational power, driven by the rapid
development of software systems directly or indirectly leveraging
artificial intelligence, as well as critical infrastructure applications
requiring the use of HPC (such as flood or tsunami simulations), the need for
accurate energy measurement tools for computing systems continues to grow.
In addition, global trends toward sustainability and responsible energy
consumption further emphasize the importance of high-quality energy profiling.

As central processing units (CPUs) represent the core components of modern
computing systems, accurate measurement of their energy consumption plays a
crucial role. Such data can be applied in multiple contexts, including the
software solutions optimization, energy-aware billing models, and performance
evaluation in HPC environments.

While Intel has long dominated the CPU market, AMD has also developed processors
targeting server and cluster workloads, providing various telemetry
capabilities including support for Intel’s RAPL (Running Average Power
Limit) interface and other AMD-specific solutions. The latter include kernel
modules and community-driven open-source tools exposing telemetry features of
AMD hardware.

This research aims to conduct a comparative analysis of energy measurement
interfaces available on AMD CPUs, focusing on the evaluation of accuracy,
resolution, and usability. Specifically, the study contrasts Intel-inspired
RAPL with AMD’s native telemetry mechanisms in order to determine their strengths,
limitations, and potential applications in modern energy-aware computing.

\input{keywords}
